---
title: "hw02-Cecilia-Li"
author: Cecilia li
output: github_document
---

**1) Create a data dictionary**

Finished data direction is saved as a md file inside the data folder of the hw02 subdirectory.

**2) Import Data in R**

```{r}
data1 <-
  read.csv(
  'nba2017-player-statistics.csv',
  colClasses = c(
  'Player' = 'character',
  'Team' = 'character',
  'Experience' = 'character',
  'Salary' = 'double'
  )
  )
  str(data1)
```
```{r}
library(readr)
```
```{r}
data2 <-
  read_csv(
  'nba2017-player-statistics.csv',
  col_types = list(
  Player = col_character(),
  Team = col_character(),
  Position = col_factor(c('C', 'PF', 'PG','SF', 'SG')),
  Experience = col_character(),
  Salary = col_double(),
  Rank = col_integer(),
  Age = col_integer(),
  GP = col_integer(),
  GS = col_integer(),
  MIN = col_integer(),
  FGM = col_integer(),
  FGA = col_integer(),
  Points3 = col_integer(),
  Points3_atts = col_integer(),
  Points2 = col_integer(),
  Points2_atts = col_integer(),
  FTM = col_integer(),
  FTA = col_integer(),
  OREB = col_integer(),
  DREB = col_integer(),
  AST = col_integer(),
  STL = col_integer(),
  STL = col_integer(),
  BLK = col_integer(),
  TO = col_integer()
  ))
str(data2)
                
```
**3) Right after importing the data**
```{r}
data1$Experience[data1$Experience == "R"] <- "1" 
Experience <- as.integer(data1$Experience)
data1
```
**4) Performance of players**
```{r}
library(dplyr)
library(ggplot2)
Missed_FG <- data1$FGA - data1$FGM
Missed_FT <- data1$FTA - data1$FTM
PTS <- (data1$Points2) * 2 + (data1$Points3) * 3 + data1$FTM
REB <- data1$DREB + data1$OREB
data1$EFF <-
(PTS + REB + data1$AST + data1$STL + data1$BLK - Missed_FG - Missed_FT - data1$TO) / data1$GP
summary(data1$EFF)
```
```{r}
hist(
  data1$EFF,
  col = 'grey',
  xlab = "EFF",
  ylab = "Frequency",
  main = 'Histogram of Efficiency (EFF)'
  )
```

```{r}
library(dplyr)
top_10 <- slice(arrange(data1, desc(EFF)), 1:10)
top_10
```

```{r}
#Negative EFF
negative <- select(filter(data1, EFF <0), Player)
negative
```

```{r}
#Correlation between EFF and PTS
cor(data1$EFF, PTS)
#Correlation between EFF and REB
cor(data1$EFF, REB)
#Correlation between EFF and AST
cor(data1$EFF, data1$AST)
#Correlation between EFF and STL
cor(data1$EFF, data1$STL)
#Correlation between EFF and BLK
cor(data1$EFF, data1$BLK)
#Correlation between EFF and Missed FG
-cor(data1$EFF, Missed_FG)
#Correlation between EFF and Missed FT
-cor(data1$EFF, Missed_FT)
#Correlation between EFF and TO
-cor(data1$EFF, data1$TO)
#Correlation between EFF and GP
cor(data1$EFF, data1$GP)
```

```{r}
#Display the computed correlations in descending order, either in a vector or a data frame. And create a barchart with the correlations.
my_table <-
  data.frame(
  col1 = c(
  0.8588644,
  0.7634501,
  0.6689232,
  0.6957286,
  0.5679571,-0.7722477
  ,-0.7271456,-0.8003289
  ),
  col2 = c('PTS', 'REB', 'AST', 'STL', 'BLK', 'Missed FG', 'Missed FT', 'TO')
  )
names(my_table) <- c('Correlations','Stats')
my_table <- arrange(my_table, desc(Correlations))
barplot(
  as.numeric(my_table$Correlations),
  ylim = c(-1, 1),
  space = c(0.2,0),
  main = 'Correlations between Player Stats and EFF',
  col = c('grey', 'grey', 'grey', 'grey', 'grey', 'coral', 'coral', 'coral'),
  names.arg = c('PTS', 'REB', 'STL', 'AST', 'BLK', 'Missed FT', 'Missed FG', 'TO'),
  cex.names = 0.6,
  border = NA
  )
  abline(h = 0)
```
**5) Efficiency and Salary**
```{r}
scatter.smooth(
  x = data1$EFF,
  y = data1$Salary,
  xlab = "Efficiency",
  ylab = "Salary",
  main = "Efficiency and Salary"
  )
  lines(lowess(data1$EFF, data1$Salary),
  lwd = 3,
  col = 'red')
  text(30,
  28000000,
  labels = 'lowess',
  cex = 1.5,
  col = 'red')
```

```{r}
#Linear correlation coefficient
cor(data1$EFF, data1$Salary)
```
According to the calculation and the graph above, the relationship between these two variables, Efficiency and Salary, are likely to be correlated. By definition, a correlation coefficient is a statistical measure of the degree to which changes to the value of one variable predict change to the value of another. With 0.66 correlation, it means that 1 unit change in efficiency approximately contributed to 0.66 change in salary. 

```{r}
#Players that have an MPG value of 20 or more minutes per game.
new_table <- filter(data1, (data1$MIN / data1$GP) >= 20)
player2 <- select(filter(data1, (data1$MIN / data1$GP) >= 20), Player)
player2
```
```{r}
new_table
```

```{r}
# Use this data frame to create a scatterplot between Efficiency and Salary, including a lowess line
scatter.smooth(x = new_table$EFF,
               y = new_table$Salary,
               main = "Scatterplot between Efficiency and Salary (MPG>20)",
               xlab = "Efficiency",
               ylab = "Salary"
                 )
               lines(lowess(new_table$EFF, new_table$Salary),
               lwd = 3,
               col = 'red')
               text(30,
               28000000,
               labels = 'lowess',
               cex = 1.5,
               col = 'red')
```

```{r}
#Linear correlation coefficient
cor(new_table$Salary, new_table$EFF)
```
_What can you say about the relationship between these two variables for the set of “more established players”_

The relationship between EFF and Salary in the group of the more established players is hard to observe, and whether or not this has a correlation is unknown. This is because that the correlation calculated from the "More Established Players" is smaller than that calculated from the whole data set.


**6) Comments and Reflections**

•What things were hard, even though you saw them in class/lab?

Applying functions was really hard, and determing the data set for usage.

• What was easy(-ish) even though we haven’t done it in class/lab?

Analyzing graphs was easy because of my prior experience in data science. 

• Did you need help to complete the assignment? If so, what kind of help?

Yes, I consulted online resources on the code formats and applications.

• How much time did it take to complete this HW?

Around 5 hours.

• What was the most time consuming part?

The most time consuming part was having to get all the code running correctly and rewriting the code every time. 

• Was there anything that you did not understand? or fully grasped?

I do not know how to vary the code based on different functions.

• Was there anything frustrating in particular?

The conclusions are very open ended, providing us with no hint to go about in certain directions.

• Was there anything exciting? Something that you feel proud of? (Don’t be shy, we
won’t tell anyone).

Learning how to download data sets and loading onto the file, the set up, and finishing the assignment was challenging but rewarding. 





